NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-31 paper/background.tex
section 2, 3 global pass

Many of the recent NLP end-to-end neural network models, despite having potentially drastically different network architectures, share a similar high-level design that consists of three distinct stages (encode, align, classify, see Fig.~\ref{fig:modelPipeline}).
%
The existence of shared conceptual structures means the proposed tool, although designed for NLI tasks, can be readily generalized for a much broader set of applications.

\begin{figure}[htbp]
\vspace{-2mm}
\centering
 \includegraphics[width=1.0\linewidth]{end2end}
 \vspace{-5mm}
 \caption{The shared structure of end-to-end NLP neural network models.}
  \vspace{-2mm}
\label{fig:modelPipeline}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.85\linewidth]{attentionIllustration}
  \vspace{-2mm}
 \caption{Attention can be naturally explained as a form of alignment which exposes an interpretable layer in end-to-end neural networks. On the left, the matrix shows the attention between between premise and hypothesis pair. On the right, attention can also be interpreted as alignment.}
\label{fig:attention}
  \vspace{-2mm}
\end{figure}

>>>>

Many of the recent NLP end-to-end neural network models, despite having potentially drastically different network architectures, share a similar high-level design that consists of three distinct stages (encode, align, classify, see Fig.~\ref{fig:end2end}).
%
The existence of shared conceptual structures means the proposed tool, although designed for NLI tasks, can be readily generalized for a much broader set of applications.

\begin{figure}[htbp]
\centering
 \includegraphics[width=1.0\linewidth]{end2end}
 \vspace{-6mm}
 \caption{The shared structure of end-to-end NLP neural network models.}
  \vspace{-5mm}
\label{fig:end2end}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.85\linewidth]{attentionIllustration}
  \vspace{-2mm}
 \caption{Attention can be naturally explained as a form of alignment which exposes an interpretable layer in end-to-end neural networks. On the left, the matrix shows the attention between the premise and hypothesis pair. On the right, we illustrate that the attention can also be interpreted as alignment.}
\label{fig:attention}
\vspace{-2mm}
\end{figure}

### 2018-03-31 paper/relatedWorks.tex
section 2, 3 global pass

\textbf{Model Interpretability.}
Before one can solve the model interpretability challenge, it worth to understand what constitutes an explanation for a machine learning model.
As discussion in~\cite{Lipton2016, Doshi-Velez2017}, there is little consensus on what interpretability is in the context of machine learning. A taxonomy is suggested~\cite{Doshi-Velez2017} to promote a more rigorous study on such a subject. In the work by Lipton~\cite{Lipton2016}, many philosophical questions related to model interpretability are examined and the study also challenged the commonly held belief that linear models are more interpretable than deep neural networks.
%
In the machine learning community, one of the most studied areas in model interpretability is the interpretation of the convolution neural network (CNN) ~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, OlahSatyanarayanJohnson2018}.
%
These methods either identify the pattern that maximally activates a part of the network (referred to as \emph{feature visualization}, i.e., what feature does the given neuron, channel, or layer captures?), or reveals what areas in the input that contribute most to the predicted label (often referred to as \emph{attribution}, i.e., attribute the prediction label in the input domain).
%
Besides the techniques designed for specific types of models, a few works~\cite{RibeiroSinghGuestrin2016, LundbergLee2017} approach the interpretability challenge from a model agnostic perspective, which makes them applicable to different applications. In the LIME~\cite{RibeiroSinghGuestrin2016} work, a simpler linear model is fitted near the prediction of interests to provide a simpler but easier to understand snapshot of the classifier near a specific input.
%
Despite being invaluable for generating a human understandable explanation, however, the lack of the ability to access and explore the internal states of the system limits their application. 
%which is vitally important for making sense of the inner works of a model, 
To avoid similar limitations, the proposed system enables direct access and manipulation of the internals of the model.

\textbf{Visual Exploration of Neural Network Models.}
Due to the effectiveness of exploring complex relationships, visualization has long been adopted for interpreting neural network models~\cite{TzengMa2005}.
Recently, the increasing demand for interpretability for deep neural network models has motivated the introduction of many visualization works that are dedicated to neural network models.
% The classification~\cite{KrausePererNg2016}
The Prospector~\cite{KrausePererNg2016} provides a model agnostic approach for analyzing and interpreting predictions via an interactive visual interface.
%
In the work~\cite{BilalJourablooYe2018} by Bilal et al., the authors analyzed classification errors in convolutional neural networks (CNN) and answered the question of whether class hierarchy is learned in the network.
%
The DeepEyes~\cite{Pezzotti2018} focused on analyzing the training process of CNN, which provides domain experts design guideline of the deep neutral network architecture.
%
The ActiVis~\cite{KahngAndrewsKalro2018} introduced an interactive system tailored for tuning industry-scale neural networks in Facebook.
Finally, Liu et al.~\cite{LiuShiCao2018} presented a visualization tool for understanding the often challenging training process of deep generative models.

>>>>

\textbf{Model Interpretability.}
Before one can solve the model interpretability challenge, it worth to understand what constitutes an explanation for a machine learning model.
As discussion in~\cite{Lipton2016, Doshi-Velez2017}, there is little consensus on what interpretability is in the context of machine learning. A taxonomy is suggested~\cite{Doshi-Velez2017} to promote a more rigorous study on such a subject. In the work by Lipton~\cite{Lipton2016}, many philosophical questions related to model interpretability are examined and the study also challenged the commonly held belief that linear models are more interpretable than deep neural networks.
%
In the machine learning community, one of the most studied areas in model interpretability is the interpretation of the convolution neural network (CNN) ~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, OlahSatyanarayanJohnson2018}.
%
These methods either identify the pattern that maximally activates a part of the network (referred to as \emph{feature visualization}, i.e., what feature does the given neuron, channel, or layer captures?), or reveals what areas in the input that contribute most to the predicted label (often referred to as \emph{attribution}, i.e., attribute the prediction label in the input domain).
%
Besides the techniques designed for specific types of models, a few works~\cite{RibeiroSinghGuestrin2016, KrausePererNg2016, LundbergLee2017} approach the interpretability challenge from a model agnostic perspective, which makes them applicable to different applications. In the LIME~\cite{RibeiroSinghGuestrin2016} work, a simpler linear model is fitted near the prediction of interests to provide a simpler but easier to understand snapshot of the classifier near a specific input.
%
Despite being invaluable for generating a human understandable explanation, however, the lack of the ability to access and explore the internal states of the system limits their application. 
%which is vitally important for making sense of the inner works of a model, 
To avoid similar limitations, the proposed system enables direct access and manipulation of the internals of the model.

\textbf{Visual Exploration of Neural Network Models.}
Due to the effectiveness of exploring complex relationships, visualization has long been adopted for interpreting neural network models~\cite{TzengMa2005}.
Recently, the increasing demand for interpretability for machine learning models has motivated the introduction of many visualization works that are dedicated to neural network models.
% The classification~\cite{KrausePererNg2016}
%The Prospector~\cite{KrausePererNg2016} provides a model agnostic approach for analyzing and interpreting predictions via an interactive visual interface.
%
In the work~\cite{BilalJourablooYe2018} by Bilal et al., the authors analyzed classification errors in convolutional neural networks (CNN) and answered the question of whether class hierarchy is learned in the network.
%
The DeepEyes~\cite{Pezzotti2018} focused on analyzing the training process of CNN, which provided domain experts design guideline of the deep neutral network architecture.
%
The ActiVis~\cite{KahngAndrewsKalro2018} introduced an interactive system tailored for tuning industry-scale neural networks in Facebook.
The TensorFlow Graph Visualizer~\cite{Wongsuphasawat2018} work designed a hierarchical visual representation to encode complex computation graphs of deep neural networks for the TensorFlow framework~\cite{Abadi2016}.
Finally, Liu et al.~\cite{LiuShiCao2018} presented a visualization tool for understanding the often challenging training process of deep generative models.
